### 2016-07-20 SPRINGER_JOURNAL/techreport/TUD-SERG-2016-018.tex
Create techreport for accepted version.
For upload to TUD-SERG-2016-018 to TU Delft preprint server.

\documentclass[letterpaper,twoside,10pt]{book}
\usepackage{se-report}

%---------------------------------------------------------------------%
% User definable options                                              %    
%---------------------------------------------------------------------%

\title{Exception Handling Bug Hazards in Android: 
Results from a Mining Study and an Exploratory Survey}

\author{Roberta Coelho, Lucas Almeida, Georgios Gousios,\\ Arie van Deursen and Christoph Treude}

\headertitle{Exception Handling Bug Hazards in Android}

\year{2016} % for report id and copyright notice
\report{018}

%\note{Accepted for publication in the Proceedings of the Working Conference
%on Reverse Engineering (WCRE), 2010, IEEE Computer Society.}

\note{This paper has been accepted for publication in \emph{Empirical Software Engineering}, 2016.\\
It is a substantially revised and extended version of 
Roberta Coelho, Lucas Almeida, Georgios Gousios, Arie van Deursen:
Unveiling Exception Handling Bug Hazards in Android Based on GitHub and Google Code Issues.
\emph{Working Conference on Mining Software Repositories} (MSR) 2015: 134-145.
}

%\copyrightmessage{%
%\copyright{} 2016 ACM. Personal use of this material is permitted. Permission from ACM must be obtained for all other uses, in any current or future media, including reprinting/republishing this material for advertising or promotional purposes, creating new collective works, for resale or redistribution to servers or lists, or reuse of any copyrighted component of this work in other works.}

\begin{document}
\maketitle

%\includepdf[trim=0 2cm 0 2cm,clip=true]{../build/pullreqs.pdf}
\includepdf[scale=0.9,clip=true]{msr_journal.pdf}

\makebackpage
\end{document}